---
title: "Data Extraction Homework KEY"
author: "Rob Wells, Ph.D."
date: "April 17, 2020"
output: html_document
---
 Instructions
Data Assignment for Thursday, April 16

    Use Tabula to scrape the the number of students from within the State of Arkansas, from out of State, and of international students for the last five years:UofA municipal bond disclose documents and enrollment trends  
    Load these tables into Google Sheets or your GitHub account 
    Import them into R. Use R Markdown
    

```{r}
library(tidyverse)
library(rio)
library(janitor)
```
Import data
```{r}
df <- rio::import("https://docs.google.com/spreadsheets/d/1VXJrRc42Q998HnI6lUA_Ja6_ToUwBTd5Hq7x2Ly3M0Q/edit?usp=sharing")
```


```{r}
UA2015_19 <- as.data.frame(t(df))
```

```{r}
UA2015_19 <- UA2015_19 %>% 
  row_to_names(row_number = 1)
UA2015_19
```
```{r}
colnames(UA2015_19)[1] <- "Year"
UA2015_19[1:4] <- lapply(UA2015_19[1:4], as.character)
glimpse(UA2015_19)
```
```{r}
library(janitor)
UA2015_19 <- janitor::clean_names(UA2015_19)
head(UA2015_19)


```


```{r}
UA2015_19$in_state <- as.numeric(gsub(",", "", UA2015_19$in_state))
UA2015_19$out_of_state <- as.numeric(gsub(",", "", UA2015_19$out_of_state))
UA2015_19$international <- as.numeric(gsub(",", "", UA2015_19$international))
UA2015_19
```
Calculate percentage change per year for in-state, out-of-state, international

```{r}
UA2015_19 <- UA2015_19 %>%
    mutate(Pct_InState = (in_state/lag(in_state)-1))
```

```{r}
UA2015_19 <- UA2015_19 %>%
    mutate(Pct_OutState = (out_of_state/lag(out_of_state)-1))
```

```{r}
UA2015_19 <- UA2015_19 %>%
    mutate(Pct_Intl = (international/lag(international)-1))
```
    Create a new total by year that adds the in-state,out-of-state and international


```{r}
UA2015_19 <- UA2015_19 %>%
   mutate(Total = (in_state + out_of_state + international))
UA2015_19
```

```{r}
UA2015_19 <- UA2015_19 %>%
   mutate(TotalPctChg = (Total/lag(Total)-1))
UA2015_19
```

```{r}
#Format into percents
library(formattable)
#
UA2015_19$TotalPctChg <- percent(UA2015_19$TotalPctChg)
UA2015_19
```


Graph percentage change over the year of this new total.
```{r}
UA2015_19 %>% 
  ggplot(aes(x = year, y = TotalPctChg, fill = TotalPctChg)) +
geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = TotalPctChg), hjust = -.5, size = 2) +
  coord_flip() +
  labs(title = "Percent Change in UofA Enrollment, 2015-2019",
       subtitle = "Total Enrollment: 27,559 in 2019",
       caption = "Graphic by Rob Wells, 4-17-2020",
       y="Source: UA Municipal Bond Disclosure",
       x="")
```

Then: Import Zillow data

    Retrieve Zillow data.  http://files.zillowstatic.com/research/public/Metro/Metro_Zri_AllHomesPlusMultifamily.csv
    Put Zillow data on Google Sheets, Import into R. Use the same R Markdown document
    Calculate percentage change per year of the Zillow index for Fayetteville only.
    Graph percentage change per year in the Zillow index

```{r}
#Import all US cities
Zillow <- rio::import('http://files.zillowstatic.com/research/public/Metro/Metro_Zri_AllHomesPlusMultifamily.csv')

```
Filter to just Fayetteville
```{r}
ZillowAR <- Zillow %>% filter(RegionID == "394590")
head(ZillowAR)
```
```{r}
#Clean Headers - Janitor package
#install.packages("janitor")
library(janitor)
ZillowAR <- janitor::clean_names(ZillowAR)
head(ZillowAR)
```
Calculate percentage change per year of the Zillow index for Fayetteville only.


First, total the years
```{r}
ZillowAR <-  ZillowAR %>%
  mutate(Yr2010 = rowSums(.[4:7])/4) %>%  
  mutate(Yr2011 = rowSums(.[8:19])/12) %>% 
  mutate(Yr2012 = rowSums(.[20:31])/12) %>% 
  mutate(Yr2013 = rowSums(.[32:43], na.rm=TRUE)/11) %>%
  mutate(Yr2014 = rowSums(.[44:55])/12) %>% 
  mutate(Yr2015 = rowSums(.[56:67])/12) %>% 
  mutate(Yr2016 = rowSums(.[68:79])/12) %>%
  mutate(Yr2017 = rowSums(.[80:91])/12) %>%
  mutate(Yr2018 = rowSums(.[92:103])/12) %>% 
  mutate(Yr2019 = rowSums(.[104:115])/12)
```

```{r}
ZillowARYR <- ZillowAR %>% 
  select(region_name, Yr2010, Yr2011, Yr2012, Yr2013, Yr2014, Yr2015, Yr2016,
         Yr2017, Yr2018, Yr2019)
ZillowARYR
```



```{r}
ZillowARYRa = as_tibble(t(ZillowARYR), rownames = "row_names")
ZillowARYRa
```
```{r}
ZillowARYR <- as.data.frame(ZillowARYRa)
ZillowARYR
```

```{r}
colnames(ZillowARYR)[1:3] <- c("x",  "Year", "ZillowIndex")

```

```{r}
#Remove rows with NA
ZillowARYR <- ZillowARYR[-c(1,2), ] 
ZillowARYR
```
```{r}
ZillowARYR[3] <- lapply(ZillowARYR[3], as.numeric)
glimpse(ZillowARYR)
```
```{r}
ZillowARYR$Year <- gsub("Yr", "", ZillowARYR$Year)
ZillowARYR
```

    Graph percentage change per year in the Zillow index
```{r}
ZillowARYR <- ZillowARYR %>%
   mutate(PctChg = (ZillowIndex/lag(ZillowIndex)-1))
ZillowARYR
```

```{r}
#Format into percents
#library(formattable)
#
ZillowARYR$PctChg <- percent(ZillowARYR$PctChg)
ZillowARYR
```

Graph percentage change over the year of this new total.
```{r}
ZillowARYR %>% 
  ggplot(aes(x = Year, y = PctChg, fill = PctChg)) +
geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = PctChg), hjust = -.5, size = 2) +
  coord_flip() +
  labs(title = "Percent Change in Fayetteville Real Estate Values, 2014-2019",
         caption = "Graphic by Rob Wells, 4-17-2020",
       y="Source: Zillow Rental Index",
       x="")
```