---
title: "Gentrification in Fayetteville"
output: html_notebook
author: "Rob Wells"
date: "4/5/2020"
output: 
  html_document:
      toc: true
      toc_float: true
      toc_depth: 4
---
 

#### Jour 405v, Jour 5003, Spring 2020    

#### Advanced Reporting and Data Analysis

#### Reporting on Homelessness in Northwest Arkansas

--------------------------------------------------------------------
# Analysis of Zillow Data for Fayetteville

Load libraries
```{r}
#install.packages("rio") 
library(tidyverse)
library(rio)
```

```{r}
#Import all US cities
Zillow <- rio::import('https://raw.githubusercontent.com/profrobwells/HomelessSP2020/master/Data/Zillow%20City_ZRI_AllhomesplusMultifamily.csv')

```
Filter to just Arkansas
```{r}
ZillowAR <- Zillow %>% filter(State == "AR")
head(ZillowAR)
```

```{r}
# Number columns
ncol(ZillowAR)
```
```{r}
# Number rows
nrow(ZillowAR)
```
```{r}
#check data types
glimpse(ZillowAR)
```

```{r}
#Clean Headers - Janitor package
#install.packages("janitor")
library(janitor)
ZillowAR <- janitor::clean_names(ZillowAR)
head(ZillowAR)
```

```{r}
#Run stats on latest November 2019 data
summary(ZillowAR$x2019_11)
```   

### Percent Change, 9/2010 to 11/2019

```{r}
Z_AR_PCT <- ZillowAR %>% 
  select(region_name, metro, x2010_09, x2019_11) %>% 
  mutate(Pct_Chg=((x2019_11-x2010_09)/(x2010_09))) 
Z_AR_PCT
```

```{r}
#To quickly format into percents, load
#install.packages("formattable")
library(formattable)
#
Z_AR_PCT$Pct_Chg <- percent(Z_AR_PCT$Pct_Chg)
Z_AR_PCT
```

```{r}
#Do some math - average value for Arkansas Sept 2010
avg2010 <- mean(Z_AR_PCT$x2010_09, na.rm=TRUE)
avg2010
```

```{r}
#Do some math - average value for Arkansas Nov 2019
avg2019 <- mean(Z_AR_PCT$x2019_11, na.rm=TRUE)
avg2019
#Average value for Arkansas, Nov. 2019= 867.8966
```

```{r}
avg <- ((avg2019-avg2010)/avg2010)*100

avg <- round(avg, 2)
avg
```

### Chart Data

```{r}
Z_AR_PCT %>% 
  filter(Pct_Chg > .20) %>% 
  ggplot(aes(x = reorder(region_name, Pct_Chg), y = Pct_Chg, fill = Pct_Chg)) +
geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = Pct_Chg), hjust = -.5, size = 2) +
  scale_y_continuous(limits=c(0, .45),labels = scales::percent) +
  coord_flip() +
  labs(title = "Arkansas Real Estate Values, 2010-2019",
       subtitle = "Statewide Average: 12.7% Increase. ",
       caption = "Graphic by Rob Wells, 4-5-2020",
       y="Source: Zillow Housing Index",
       x="")
```
###  Import Enrollment Data
Assignment to use Tabula scrape data from Fall Semester Enrollment, 2010-2019 
https://learn.uark.edu/bbcswebdav/pid-8109956-dt-content-rid-92421291_1/xid-92421291_1

```{r}
x <- rio::import('https://raw.githubusercontent.com/profrobwells/HomelessSP2020/master/Data/tabula-UofA%20annual%20report%202019.csv')
```

```{r}
UA2010_19 <- as.data.frame(t(x))
```

```{r}
UA2010_19 <- UA2010_19 %>% 
  row_to_names(row_number = 1)
UA2010_19
```
```{r}
colnames(UA2010_19)[1] <- "Year"
UA2010_19[1:6] <- lapply(UA2010_19[1:6], as.character)
glimpse(UA2010_19)
```
```{r}
UA2010_19$Total <- as.numeric(gsub(",", "", UA2010_19$Total))
UA2010_19$Undergraduate <- as.numeric(gsub(",", "", UA2010_19$Undergraduate))
UA2010_19$Law <- as.numeric(gsub(",", "", UA2010_19$Law))
UA2010_19$Graduate <- as.numeric(gsub(",", "", UA2010_19$Graduate))
UA2010_19$`New Freshmen` <- as.numeric(gsub(",", "", UA2010_19$`New Freshmen`))
UA2010_19
```

```{r}
UA2010_19 <- UA2010_19 %>% 
 mutate(total_pct_change = (Total/lag(Total)-1)) 
#UA2010_19$total_pct_change <- round(UA2010_19$total_pct_change, 2)
UA2010_19
```
```{r}
#Format into percents
#library(formattable)
#
UA2010_19$total_pct_change <- percent(UA2010_19$total_pct_change)
UA2010_19
```
### Chart Data

```{r}
UA2010_19 %>% 
  ggplot(aes(x = Year, y = total_pct_change, fill = total_pct_change)) +
geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = total_pct_change), hjust = -.5, size = 2) +
  #scale_y_continuous(limits=c(0, .10),labels = scales::percent) +
  coord_flip() +
  labs(title = "Percent Change in UofA Enrollment, 2010-2019",
       subtitle = "Total Enrollment: 27,559 in 2019",
       caption = "Graphic by Rob Wells, 4-15-2020",
       y="Source: UA Annual Report",
       x="")
```

###Align With Zillow Data for Fayetteville

Subset Fayetteville Zillow data
ZillowAR <- Zillow %>% filter(State == "AR")
```{r}
FayZillow <- ZillowAR %>% 
  filter(region_id == "31543")
```

Average by years
 mutate(total_pct_change = (Total/lag(Total)-1)) 

```{r}
write_csv(FayZillow, "FayZillow.csv")

```

 
```{r}
z1 <-  FayZillow %>% 
  mutate(Yr2010 = (x2010_09 + x2010_10 + x2010_11 + x2010_12)/4) 
```

```{r}
FayZillow1 <-  FayZillow %>% 
  mutate(Yr2010 = rowSums(.[7:10])/4) %>%  
  mutate(Yr2011 = rowSums(.[11:22])/12) %>% 
  mutate(Yr2012 = rowSums(.[23:34])/12) %>% 
  mutate(Yr2013 = rowSums(.[35:46])/12) %>%
  mutate(Yr2014 = rowSums(.[47:58])/12) %>% 
  mutate(Yr2015 = rowSums(.[59:70])/12) %>% 
  mutate(Yr2016 = rowSums(.[71:82])/12) %>%
  mutate(Yr2017 = rowSums(.[83:94])/12) %>%
  mutate(Yr2018 = rowSums(.[95:106])/12) %>% 
  mutate(Yr2019 = rowSums(.[107:117])/11)
```


problem with 2013
```{r}
z2 <-  FayZillow %>% 
  mutate(Yr2013 = rowSums(.[35:46], na.rm=TRUE)/11)
```

```{r}
FayZillow1 <-  FayZillow %>% 
  mutate(Yr2010 = rowSums(.[7:10])/4) %>%  
  mutate(Yr2011 = rowSums(.[11:22])/12) %>% 
  mutate(Yr2012 = rowSums(.[23:34])/12) %>% 
  mutate(Yr2013 = rowSums(.[35:46], na.rm=TRUE)/11) %>%
  mutate(Yr2014 = rowSums(.[47:58])/12) %>% 
  mutate(Yr2015 = rowSums(.[59:70])/12) %>% 
  mutate(Yr2016 = rowSums(.[71:82])/12) %>%
  mutate(Yr2017 = rowSums(.[83:94])/12) %>%
  mutate(Yr2018 = rowSums(.[95:106])/12) %>% 
  mutate(Yr2019 = rowSums(.[107:117])/11)
```

```{r}
FayZillowYR <- FayZillow1 %>% 
  select(region_name, Yr2010, Yr2011, Yr2012, Yr2013, Yr2014, Yr2015, Yr2016,
         Yr2017, Yr2018, Yr2019)
FayZillowYR
```

```{r}
FayZillowYR = as_tibble(t(FayZillowYR), rownames = "row_names")

```

```{r}
FayZillowYR <- as.data.frame(FayZillowYR)
FayZillowYR
```
```{r}
FayZillowYR <- FayZillowYR %>% 
  row_to_names(row_number = 1)
FayZillowYR
```
```{r}
colnames(FayZillowYR)[1:2] <- c("Year", "ZillowIndex")
FayZillowYR[2] <- lapply(FayZillowYR[2], as.numeric)
glimpse(FayZillowYR)
```
```{r}
FayZillowYR$Year <- gsub("Yr", "", FayZillowYR$Year)
FayZillowYR
```

```{r}
UA_Zillow <- UA2010_19 %>% 
  inner_join(FayZillowYR, by = ("Year"))
UA_Zillow
```

```{r}
UA_Zillow <- UA_Zillow %>% 
 mutate(Zillow_pct_change = (ZillowIndex/lag(ZillowIndex)-1)) 
UA_Zillow
```
```{r}
#Format into percents
#library(formattable)
#
UA_Zillow$Zillow_pct_change <- percent(UA_Zillow$Zillow_pct_change)
UA_Zillow
```



```{r}
write_csv(UA_Zillow, "UA_Zillow.csv")

```

### Chart two items
solution: https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales

Convert to dates
```{r}
UA_Zillow$Year1 <- as.numeric(UA_Zillow$Year)

UA_Zillow$Year1 <- as.Date(as.character(UA_Zillow$Year),format="%Y")
#UA_Zillow$Year1 <- as.date(UA_Zillow$Year1)
#UA_Zillow$Year1 <- year(UA_Zillow$Year1)
#UA_Zillow$Year1 <- as.Date(UA_Zillow$Year1, origin="2010")
UA_Zillow
```

### Dual Axis Chart    

    -- First, UA Enrollment, as a pink bar
    -- Second, Zillow index, as a blue line  
    -- UA Enrollment is larger than the Zillow Index. It throws off the chart
    -- We need to adjust it so both items are reflected on the same chart. Dual axis chart
    -- So we adjust the Zillow index - multiply it by .20 - to better fit with the UA Enrollment graphic
    -- We adjust the Zillow index scale by .235 to reflect the Zillow index, not the UA Enrollment


```{r}

chart <- ggplot() + 
  geom_bar(mapping = aes(x = UA_Zillow$Year1, y = UA_Zillow$Total), stat = "identity", fill = "pink") +
  geom_line(mapping = aes(x = UA_Zillow$Year1, y = UA_Zillow$ZillowIndex*20), size = 2, color = "blue", fill = "blue") + 
  scale_x_date(name = "Year") +
  scale_y_continuous(name = "UA Enrollment", 
                     sec.axis = sec_axis(~. /23.5, name = "Zillow Index")) + 
   theme(
    axis.title.y = element_text(color = "pink"),
    axis.title.y.right = element_text(color = "blue")) +
  labs(title = "UA Enrollment, Fayetteville Real Estate Prices, 2010-2019",
       subtitle = "Total Enrollment: 27,559 in 2019",
       caption = "Graphic by Rob Wells, 4-15-2020")
chart
```

```{r}
ggsave(chart, file="chart.png",scale=2)
```    
