---
title: "ggplot cookbook"
author: "Rob Wells"
date: "4/26/2020"
output: html_document
---

###A cookbook for formatting decent graphics in ggplot
Load libraries
```{r}
#install.packages("rio") 
library(tidyverse)
library(rio)
```

```{r}
#Import data
Homeless2018 <- rio::import('https://github.com/profrobwells/HomelessSP2020/raw/master/Data/Homeless2018.csv')

```

glimpse
```{r}
glimpse(Homeless2018)
```

#Basic graphic - labels
```{r}
The15 <- Homeless2018 %>% 
  filter(district_percent_homeless > .15) %>% 
  ggplot(aes(x = reorder(district_bak, district_percent_homeless), 
             y = district_percent_homeless,
             fill = district_percent_homeless)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  #label formatting. Scales, into percentages. hjust moves to the grid
    geom_text(aes(label = scales::percent(district_percent_homeless)), position = position_stack(vjust = .5), hjust = -5., size = 2.5)  +
  #format the x axis. sets the grid to maximum 30%
    scale_y_continuous(limits=c(0, .3),labels = scales::percent) +
  coord_flip() +
  labs(title = "Homeless Children in Arkansas, 2018",
       subtitle = "Districts with More Than 15% Homeless",
       caption = "Graphic by Rob Wells, 4-26-2020",
       y="Statewide Average: 3.6%. Source: Arkansas Dept of Education",
       x="")
The15
```

# Export to hi-res file
```{r}

ggsave("Test.png",device = "png",width=9,height=6, dpi=800)
```
#Make two plots, put on one chart

#plot Benton, Springdale, Fayetteville

```{r}
NWA <- Homeless2018 %>% 
  filter(district_bak =="BENTONVILLE" | district_bak =="FAYETTEVILLE" | district_bak =="SPRINGDALE") %>% 
  ggplot(aes(x = reorder(district_bak, district_percent_homeless), 
             y = district_percent_homeless,
             fill = district_percent_homeless)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  #label formatting. Scales, into percentages. hjust moves to the grid
    geom_text(aes(label = scales::percent(district_percent_homeless)), position = position_stack(vjust = .7), hjust = -5., size = 2.5)  +
  #format the x axis. sets the grid to maximum 30%
    scale_y_continuous(limits=c(0, .03),labels = scales::percent) +
  coord_flip() +
  labs(title = "NWA Homeless Children, 2018",
       subtitle = "Statewide Average: 3.6%",
       caption = "Graphic by Rob Wells, 4-26-2020",
       y="Source: Arkansas Dept of Education",
       x="")
NWA
```

#Put two plots on one graphic. Stacked horizontally by manipulating the ncol and nrow

```{r}
#install.packages("ggpubr")
library(ggpubr)
final <- ggarrange(The15, NWA, ncol = 1, nrow = 2)
final
```


You can work on the formatting from here:
#sample code
#Adjust the size of the x and y axis fonts
theme(axis.text.x = element_text(angle = 90, hjust = 1, size=8)) +
  theme(axis.title.y = element_text(size = 8)) +
  
#A post on how to arrange multiple plots on a page:
http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/
  theme(axis.title.x = element_text(size = 8)) +
 
Stacked charts
https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html
 
#figure a state average

```{r echo=FALSE}
library(formattable)
percent(mean(Homeless2018$district_percent_homeless, na.rm=TRUE))

```

glimpse
```{r}
glimpse(Homeless2018)
```

```{r}
Homeless2018 %>%
  ggplot(aes(x = district_percent_homeless, y = reorder(district_bak, district_percent_homeless), color = district_percent_homeless < .0359))+   
  geom_point() +
  theme(axis.text.y = element_text(size=2)) 
```

```{r}
Homeless2018 %>%
  ggplot(aes(y = district_bak, x = district_percent_homeless, color = district_percent_homeless < .0359)) +   
  geom_point() +
  theme(axis.text.x = element_text(size=5)) +
   theme(axis.text.y = element_text(size=.000001)) 
```


#Scatterplot cleaning up the messy names
  axis.ticks.x=element_blank())
  theme(#axis.title.x=element_blank(),
  axis.text.x=element_blank(),
  axis.ticks.x=element_blank())


```{r}
scatter <- ggplot(Homeless2018, aes(x = district_bak, y = district_percent_homeless,  color = district_percent_homeless < .0359)) +
   geom_point(alpha = 0.3) +
   geom_smooth(method = lm) +
   theme(axis.text.x=element_blank()) +
    scale_y_continuous(limits=c(0, .4),labels = scales::percent) +
       labs(title = "Homeless Children in Arkansas, 2018",
       caption = "Graphic by Rob Wells, 4-26-2020",
       y="Percent Homeless by School District",
       x="")
scatter
```

#Formatting options explained well
https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/#remove-legend
https://stackoverflow.com/questions/23161897/how-to-change-labels-legends-in-ggplot
```{r}
scatter2 <- scatter + #labs(color='Above / Below State Average') 
 scale_color_manual(name="Above / Below State Average",
                       labels=c("Above Avg","Below Avg", "NA"),
                       values=c("red","green","gray"))
scatter2
```

# Export to hi-res file
```{r}

ggsave("scatter.png",device = "png",width=10,height=8, dpi=800)
```

```{r}
Top10<- Homeless2018 %>% 
  select(district_bak, district_percent_homeless) %>% 
   top_n(10, district_percent_homeless)
Top10
```
```{r}
colnames(Top10)[1:2] <- c("District", "Pct_Homeless")
library(formattable)
Top10$Pct_Homeless <- percent(Top10$Pct_Homeless)
```

```{r}
glimpse(Top10)
```
```{r}
Top10 %>%
  ggplot(aes(y = District, x = Pct_Homeless, color = Pct_Homeless)) +   
  geom_point() 

```


```{r}
#Formatting
#install.packages("kableExtra")
library(kableExtra)
# This makes kables
x<- Top10 %>% 
  kable() %>%
  kable_styling("striped")
```

#Put two plots on one graphic. Stacked horizontally by manipulating the ncol and nrow

```{r}
#install.packages("ggpubr")
library(ggpubr)
scatter2 <- ggarrange(scatter2, x, ncol = 1, nrow = 2)
scatter2
```
