---
title: "Homeless Children in Arkansas"
author: "Rob Wells, Ph.D."
date: "3/28/2020"
output: 
  html_document:
      toc: true
      toc_float: true
      toc_depth: 4
---
 
## Reporting on Homelessness: Data Analysis for Journalists   

#### Jour 405v, Jour 5003, Spring 2020    

--------------------------------------------------------------------
# Part 1: Homeless Children Data
Load tidyverse, janitor, lubidate
```{r include=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```
Load Data
```{r include=FALSE}
ChildrenAR <- rio::import("https://github.com/profrobwells/HomelessSP2020/blob/master/Data/2018-2019-demographic-data-for-districts.xlsx?raw=true", which = "Data", skip=8)
```
Clean up column names to they are R friendly
```{r}
ChildrenAR <- janitor::clean_names(ChildrenAR)
```

- **Run a summary of the homeless data.** 

What is the median homeless percentage?   
What is the average (mean)?   
What is the maximum?   

```{r}
summary(ChildrenAR$district_percent_homeless)
```

- **Create a table with all school districts with more than 10% homeless**

HINT: SEE THE R_Intro_Jan15, the code for "Build a table with all counties above that median income"

```{r}
  ChildrenAR %>% 
  select(district_name, district_percent_homeless) %>% 
  filter(district_percent_homeless >= .10) %>% 
  arrange(desc(district_percent_homeless))

```
Create Children10 table, from code above, table with all school districts with more than 10% homeless
```{r}
Children10 <- ChildrenAR %>% 
  select(district_name, district_percent_homeless) %>% 
  filter(district_percent_homeless >= .10) %>% 
  arrange(desc(district_percent_homeless))
```

## Change Over Time

For your reference!
https://profrobwells.github.io/HomelessSP2020/SF_311_Calls_UofA.html

Import 2013-14 homeless data, call it ChildrenAR2013, clean the labels

```{r}
ChildrenAR2013 <- rio::import("https://github.com/profrobwells/HomelessSP2020/blob/master/Data/2013-2014-demographics-data-for-districts-2.xlsx?raw=true", which = "Data", skip=8)
# Clean up column names to they are R friendly
ChildrenAR2013 <- janitor::clean_names(ChildrenAR2013)
```

- **Create a table with just the district name and the homeless data for 2018-19**      

Call it table2018
```{r}
table2018 <- ChildrenAR %>% 
  select(district_name, district_percent_homeless, district_lea) %>% 
  arrange(desc(district_percent_homeless))
```

Rename a column name
```{r}
colnames(Children10)[2] <- "Pct_Homeless_2018"
#It takes the second column in Children10 and assigns the new name #"Pct_Homeless_2018"
names(Children10)
```
- **Rename the appropriate table2018 column to "Pct_Homeless_2018" ** 
```{r}
colnames(table2018)[2] <- "Pct_Homeless_2018"
```
- **Follow the same steps for the 2013 data:** 
Create table2013 
Rename the appropriate table2013 column to "Pct_Homeless_2013"
```{r}
table2013 <- ChildrenAR2013 %>% 
  select(district_name, district_percent_homeless,  district_lea) %>% 
  arrange(desc(district_percent_homeless))
#
colnames(table2013)[2] <- "Pct_Homeless_2013"
#
#convert all text to lowercase
#table2013$district_name <- tolower(table2013$district_name)
```
- **Joining Tables**   
```{r}
newtable <- table2018 %>% 
  inner_join(table2013, by=("district_lea"))
```
This matches the tables by name. They are joined by values in district_name   

- **How many records were left out?**    
8 Records.   
Use anti_join to determine the 2013 records not in 2018 
```{r}
table2013 %>% 
  anti_join(table2018, by=("district_lea"))
```

Making a table with the same thing above
```{r}
leftout <- table2013 %>% 
  anti_join(table2018, by=("district_lea"))
```

- **Math!**  

This creates a new column with the percentage point difference
```{r}
newtable <- newtable %>% 
  mutate(difference = Pct_Homeless_2018-Pct_Homeless_2013)
```

- **Produce a table with top 20 schools with the greatest increase in student homelessness since 2013, measured by percentage point difference**  

```{r}
top10worst <-  newtable %>% 
  top_n(20, difference) %>% 
  arrange(desc(difference))
```

# Basic Map Exercise
- **We will create a map of homeless children by schools in Arkansas**  

This lesson references the Machlis Ch 11 map exercise as a template. 
It also references Andrew Ba Tran's map tutorial
https://github.com/profrobwells/HomelessSP2020/blob/master/Readings/Tran%20-%20Geolocating.pdf

#### HERE WE GO!

Use ChildrenAR
Join ChildrenAR with addresses

Load Addresses
```{r}
School_Address <- rio::import("https://github.com/profrobwells/HomelessSP2020/blob/master/Data/SDI%20Districts%20Printed-02-26-2020-11_38_05.xlsx?raw=true", skip =1)
```

Convert School address LEA to numeric and match
```{r}
School_Address$LEA <- as.numeric(School_Address$LEA)
ChildrenAR$LEA <- ChildrenAR$district_lea
```
Now we join School_Address to ChildrenAR

```{r}
abc <- School_Address %>% 
  select(LEA, Name, Address) %>% 
  inner_join(ChildrenAR, by="LEA")
```

#QUESTION: How many records in new combined dataframe? What is the difference?

```{r}
abc_not <- School_Address %>% 
  select(LEA, Name, Address) %>% 
  anti_join(ChildrenAR, by="LEA")
```

Df with 261. which ones dropped? Lost two

NEXT, we geocode the addresses using mutate_geocode
This is a little involved... 

What I did:
?register_google
Google API Key https://www.youtube.com/watch?v=1JNwpp5L4vM
https://console.cloud.google.com/apis/credentials?folder=&organizationId=&project=phonic-botany-269820
register_google(key = "MY TOP SECRET KEY")
schools_geocoded <- mutate_geocode(abc, Address)
write.csv(schools_geocoded, "schools_geocoded.csv")

- **Shortcut!!!**

```{r include=FALSE}
schools_geocoded <- rio::import("https://raw.githubusercontent.com/profrobwells/HomelessSP2020/master/Data/schools_geocoded.csv")
```

Next: Simplify the school district names

```{r}
schools_geocoded$Name_Cleaned <- gsub("SCHOOL", "", schools_geocoded$Name)
```

```{r}
schools_geocoded$Name_Cleaned <- gsub("DISTRICT", "", schools_geocoded$Name)
```

```{r}
schools_geocoded$Name_Cleaned <- gsub("SCHOOL", "", schools_geocoded$Name_Cleaned) 
```
Load map town shapes for Arkansas with Tigris.

If you don't have tigris or ggplot2 or sf installed yet, uncomment and run the line below
```{r include=FALSE}
#install.packages("tigris", "sf", "ggplot2")
library(tigris) 
library(sf) 
library(ggplot2)
```
```{r}
# set sf option
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

And we finally load the school district boundary map
```{r include=FALSE}
#Uncomment if you want to run this
ar_schools <- school_districts("05")
```

Basic Map of School Districts
```{r}
ggplot(ar_schools) +
  geom_sf() +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent')) + labs(title="AR School Districts")
```


Adding layer with homeless data   

```{r}
set.seed(25)
```
We create a temporary table "bf" that has the district school names, percent homeless, the lon, lat and filtered for districts with more that 10% homeless

```{r}
bf<- schools_geocoded %>% 
  select(district_name, Name_Cleaned, district_percent_homeless, lon, lat) %>% 
  filter(district_percent_homeless >= .10)
```

Plot map with subset data

```{r}
ggplot(ar_schools) +
  geom_sf() +
  geom_point(data=bf, aes(x=lon, y=lat, size=district_percent_homeless, color=district_percent_homeless), show.legend = FALSE) + 
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent')) + 
  labs(title="Homeless Children in Arkansas",
       subtitle = " School Districts With More Than 10% Homeless Students",
       caption = "Source: Ar Dept of Education. Graphic by Wells. 3-1-2020")
```


Clean it up!   

```{r}
bf <- janitor::clean_names(bf)
```

Plot with labels
```{r}
library(ggrepel)
```

```{r}
ggplot(ar_schools) +
  geom_sf() +
  geom_point(data=bf, aes(x=lon, y=lat, 
                          size=district_percent_homeless, color=district_percent_homeless), 
             show.legend = FALSE) + 
  geom_label_repel(data=bf, aes(x=lon, y=lat, label = name_cleaned), size=2.5, color="blue") +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent')) + 
  labs(title="Homeless Children in Arkansas",
       subtitle = " School Districts With More Than 10% Homeless Students",
       caption = "Source: Ar Dept of Education. Graphic by Wells. 3-1-2020")
```

<br>    

- **Task:Create a dataframe with the top 10 districts with the largest
percentage increase in homelessness since 2013 and map it.**   


# Homeless Children Data Analysis
#March 14, 2019

In this section, we analyze homeless children data over time and compare previous years of homeless children data.
--Take the 2018-2019 data and compare to the 2014-2015 data
--Following the same workflow in the Census exercise, import the two datasets, 
rename the columns, clean the district names, calculate percentage change in homeless
1) --Produce a table with statewide changes in homelessness and the special needs index from 2018-2019 vs 2014-2015. 
 2) --Produce a smaller chart with all of the districts you have contacted 

## Part 1: import the two datasets, rename the columns, clean the district names

ChildrenAR is already loaded from above. Rename as ChildAR2019

```{r}
ChildAR2019 <- ChildrenAR
```
Rename columns anticipating merge with older data

```{r}
colnames(ChildAR2019)[3] <- c("enroll_2019")
colnames(ChildAR2019)[5:6] <- c("special_ed_pct_2019", "homeless_pct_2019")
```

Load ChildrenAR 2014-15 data, clean and rename
```{r}
ChildAR2015 <- rio::import("https://github.com/profrobwells/HomelessSP2020/blob/master/Data/2014-2015-demographics-for-districts.xlsx?raw=true", which = "Data", skip=8)
ChildAR2015 <- janitor::clean_names(ChildAR2015)
colnames(ChildAR2015)[3] <- c("enroll_2015")
colnames(ChildAR2015)[5:6] <- c("special_ed_pct_2015", "homeless_pct_2015")
```

SELECT SUBSET WITH DISTRICT NAME POVERTY SPECIAL NEEDS ENROLLMENT AND MERGE
Merge Census Data

```{r}
temp <- ChildAR2019 %>%
  inner_join(ChildAR2015, by=("district_lea"))
```
Boil it down

```{r}
Child2019_15 <- temp %>% 
  select(district_name.x, enroll_2019, enroll_2015, special_ed_pct_2019, special_ed_pct_2015,
                       homeless_pct_2019, homeless_pct_2015, district_lea)
colnames(Child2019_15)[1] <- c("district")
```

Simplify the school district names
Backup distict name
```{r}
Child2019_15$district_bak <- Child2019_15$district
Child2019_15$district <- gsub("SCHOOL", "", Child2019_15$district) 
Child2019_15$district  <- gsub("DISTRICT", "", Child2019_15$district)
Child2019_15$district <- gsub("SCHOOL", "", Child2019_15$district) 
Child2019_15$district <- gsub(" ", "", Child2019_15$district) 
```
Part 2: Calculate calculate percentage change in homeless: How has homeless changed from 2015-19?

```{r}
Child2019_15 <- Child2019_15 %>%
  mutate(homeless_pct_diff = (homeless_pct_2019-homeless_pct_2015)/homeless_pct_2015) %>%
  arrange(desc(homeless_pct_diff))
```

```{r}
library(formattable)

Child2019_15$homeless_pct_diff <- percent(Child2019_15$homeless_pct_diff)
```
How has special needs changed from 2015-19

```{r}
Child2019_15 <- Child2019_15 %>%
  mutate(special_pct_diff = (special_ed_pct_2019-special_ed_pct_2015)/special_ed_pct_2015) %>%
  arrange(desc(special_pct_diff))

library(formattable)

Child2019_15$special_pct_diff <- percent(Child2019_15$special_pct_diff)
```

--Produce a table with statewide changes in homelessness and the special needs index from 2018-2019 vs 2014-1025. 

```{r}
Child2019_15 %>% 
  select(district, homeless_pct_diff, special_pct_diff)
```

Save to a table
```{r}
Child2019_15_summary <- Child2019_15 %>% 
  select(district, homeless_pct_diff, special_pct_diff)
```

Your schools
```{r}
Child2019_15 %>% 
filter(district=="BERRYVILLE" | district=="JASPER" | district=="LEADHILL")
```

## Notes Beyond this point

install.packages("ggpubr")
library("ggpubr")


cor(x, y, method = c("pearson", "kendall", "spearman"))
cor.test(x, y, method=c("pearson", "kendall", "spearman"))

res <- cor.test(Child2019_15$homeless_pct_2019, Child2019_15$special_ed_pct_2019, 
                    method = "pearson")
                    
res

res1<- cor.test(Child2019_15$, Child2019_15$special_pct_diff, 
                    method = "pearson")
