---
title: "Wisconsin Evictions Analysis"
author: "Rob Wells"
date: "8/4/2020"
output: html_document
---

# Analyzing Eviction Court Data From Milwaukee
### Data analysis by Rob Wells


- **The records were scraped from this website**  
https://wcca.wicourts.gov/advanced.html 

https://github.com/Howard-Center-Investigations/evictions/blob/master/cleaning/output/all_standardized/sta_wi_milwaukee_2019_2020.csv

--------------------------------------------------------------------
- **Part 1: Load Libraries and Data**

```{r include=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
library(kableExtra)
```
Load Data for Evictions tab
```{r}
milw <- rio::import("https://raw.githubusercontent.com/Howard-Center-Investigations/evictions/master/cleaning/output/cities_original/wi_milwaukee_2019_2020.csv?token=AK7LSTECLG4CSE7CSOGT7AS7FRWA2") 
```




- **Clean names**
```{r}
milw <- janitor::clean_names(milw)
```
- **Check data fields**
```{r}
glimpse(milw)
```
- **Process dates** 
```{r}
#Process dates
milw$date <- ymd(milw$file_date)
```

```{r}
milw <- milw %>% 
  mutate(yearmon = format(date, "%Y-%m"))
```
- **Check data fields**
```{r}
glimpse(milw)
```
### Total Eviction Filings by Month in Wisconsin through July 31, 2020
**Wisconsin eviction moratorium was March 27 - May 27, 2020**
```{r}
z <- milw %>% 
  group_by(yearmon) %>%
  count(yearmon)
z
#write.csv(z, "miltotal.8.4.2020.csv")
```

- **Chart of Wisconsin Eviction Filings in 2020**  
**Wisconsin eviction moratorium was March 27 - May 27, 2020**
```{r}

ggplot(z, aes(x=yearmon, y=n, fill=yearmon)) +
  geom_col() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Wisconsin Eviction Filings Per Month Jan. 2019-July 2020",y="Evictions", x= "", 
       caption = "Graphic by Rob Wells. 8-4-2020", subtitle = "Source: Wisconsin Circuit Court records")+
  theme(legend.position="none") 
ggsave("Wisc_Evict_8_4_2020.png",device = "png",width=9,height=6, dpi=500)
```

-- **Clean the columns - remove stray spaces**

```{r}
p <- milw
```


p$plaintiff <- str_trim(p$plaintiff)
p$defendant <- str_trim(p$defendant)
p$plaintiff <- gsub("[[:punct:]]", "", p$plaintiff)
p$defendant <- gsub("[[:punct:]]", "", p$defendant)
head(p)
#x$plaintiff <- gsub(" ", "_", x$plaintiff)


-- **Summary of Milwaukee Plaintiffs **
```{r}
#Build a simple summary table
plaintiff1 <- p %>% 
  select(plaintiff_one, defendant_one, date) %>% 
         group_by(plaintiff_one) %>% 
         count(plaintiff_one) %>% 
         ungroup() %>% 
         arrange(desc(n))

#write.csv(plaintiff1, "milplaintiff.7.6.2020.csv")


plaintiff1
```

--**Context**: During the entire course of this study, the 6,076 eviction records since January 2019, the average number of evictions by a landlord was 3. 
```{r}
summary(plaintiff1)
```


-- **Milwaukee All since March 27 2020**
```{r}
x <- p %>% 
  filter(date > '2020-03-27')

write.csv(x, "mil_all_march27forward.csv")

```


-- **Milwaukee evictions / Plaintiffs since March 27 2020**
```{r}
moratorium <- p %>% 
  filter(date > '2020-03-27') %>% 
  select(plaintiff_one, defendant_one, date) %>% 
         group_by(plaintiff_one) %>% 
         count(plaintiff_one) %>% 
         ungroup() %>% 
         arrange(desc(n))

#write.csv(moratorium, "milplaintiff.8.4.2020.csv")

```

-- **Milwaukee Plaintiffs since March 27 2020**
```{r}
top <- p %>% 
  filter(date > '2020-03-27') %>% 
  select(plaintiff_one, yearmon) %>% 
         group_by(plaintiff_one, yearmon) %>% 
         count(plaintiff_one) %>% 
         filter(n > 5) %>% 
         ungroup() %>% 
         arrange(yearmon) %>% 
         arrange(desc(n))
write.csv(top, "mil_evictors_mo_8_4_2020.csv")
```

--**Context**: Since the moratorium, the 1,406 eviction records since March 27, 2020, the average number of evictions by a landlord was 2. 
Since the moratorium started in March 27, 2020, there have been 1,406 evictions. Berrada alone (checking the other entities soon) had 306 evictions, or 22% of the total. One of every five evictions was filed by Berrada.  On average, a landlord would have evicted just three people. Berrada evicted 306.
```{r}
summary(moratorium)
```

##Radical Cleaning of Plaintiffs

```{r}
p2 <- p

p2$plaintiff_fixed <- p2$plaintiff_one
```


```{r}
glimpse(p2)
```


--Clean plaintiff_fixed

```{r}
p2$plaintiff_fixed <- str_remove(str_replace_all(str_remove(p2$plaintiff_fixed, "^.*\\n"), "\\n", " "), "\\|.*")
```

```{r}
p2$plaintiff_fixed <- str_trim(p2$plaintiff_fixed)
p2$plaintiff_fixed <- tolower(p2$plaintiff_fixed)
```

--Clean plaintiff_one

```{r}
p2$plaintiff_one <- str_remove(str_replace_all(str_remove(p2$plaintiff_one, "^.*\\n"), "\\n", " "), "\\|.*")
```

```{r}
p2$plaintiff_one <- str_trim(p2$plaintiff_one)
p2$plaintiff_one <- tolower(p2$plaintiff_one)
```

--**Rename and Reclassify Landlords, Substitute top names and normalize**
```{r}
p2 <- p2 %>% 
  mutate(plaintiff_fixed = case_when(plaintiff_one=="highland llc" ~ "berrada",
                                     plaintiff_one=="meaux park llc" ~ "berrada",
                                     plaintiff_one=="spring court apartments llc" ~ "berrada",
                                     plaintiff_one=="dineen park townhomes llc" ~ "berrada",
                                     plaintiff_one=="arbor gardens apartments llc" ~ "berrada",
                                     plaintiff_one=="luther haven llc" ~ "berrada",
                                     plaintiff_one=="willow court apartments llc" ~ "berrada",
                                     plaintiff_one=="garden terrace homes llc" ~ "berrada",
                                     plaintiff_one=="good hope townhomes ii llc" ~ "berrada",
                                     plaintiff_one=="good hope townhomes llc" ~ "berrada",
                                     plaintiff_one=="greenbrook place apartments llc" ~ "berrada",
                                     plaintiff_one=="colonial pointe ii apartments llc" ~ "berrada",
                                     plaintiff_one=="colonial pointe il apartments llc" ~ "berrada",
                                     plaintiff_one=="colonial pointe i apartments llc" ~ "berrada",
                                     plaintiff_one=="mariners pointe apartments llc" ~ "berrada",
                                     plaintiff_one=="berrada properties llc" ~ "berrada",
                                     plaintiff_one=="yb florist llc" ~ "berrada",
                                     plaintiff_one=="yb florist,llc" ~ "berrada",
                                     plaintiff_one=="yb florist, llc" ~ "berrada",
                                     plaintiff_one=="yb medford llc" ~ "berrada",
                                     plaintiff_one=="yb silver spring llc" ~ "berrada",
                                     plaintiff_one=="yb victory llc" ~ "berrada",
                                     plaintiff_one=="yb victory llc" ~ "berrada",
                                     plaintiff_one=="paul e hurst, co berrada properties" ~ "berrada",
                                     plaintiff_one=="falcon crest llc" ~ "berrada",
                                     plaintiff_one=="falcon crest llc" ~ "berrada",
                                     plaintiff_one=="	wisniewski, michael s" ~ "michael_s_wisniewski",
                                     plaintiff_one=="bruckner, daniel w" ~ "daniel_w_bruckner",
                                     plaintiff_one=="	nawrocki, gregory w" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="nawrocki, gregory" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="nawrocki, gragory w" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="nawriocki, gregory w" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="nawrocki, gregory w" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="martin, bruce" ~ "bruce_martin",
                                     plaintiff_one=="housing authority of the city of milwaukee" ~ "milwaukee_city_housing_authority",
                                     plaintiff_one=="city of milwaukee" ~ "milwaukee_city",
                                     plaintiff_one=="city  of milwaukee" ~ "milwaukee_city",
                                     plaintiff_one=="milwaukee living llc" ~ "milwaukee_living",
                                     plaintiff_one=="milwaukee capital properties llc" ~ "milwaukee_capital_properties",
                                     plaintiff_one=="good hope townhomes ii llc" ~ "good_hope_townhomes",
                                     plaintiff_one=="good hope townhomes llc" ~ "good_hope_townhomes",
                                     plaintiff_one=="howard and th street holding company" ~ "howard_76_st_holding_co",
                                     plaintiff_one=="maas, robert" ~ "robert_maas",
                                     plaintiff_one=="maas, robert j" ~ "robert_maas",
                                    TRUE~plaintiff_one)) 

```


-- **Summary of Cleaned Milwaukee Plaintiffs **
```{r}
#Build a simple summary table
plaintiff2 <- p2 %>% 
  select(plaintiff_fixed, date) %>% 
         group_by(plaintiff_fixed) %>% 
         count(plaintiff_fixed) %>% 
         ungroup() %>% 
         arrange(desc(n))

write.csv(plaintiff2, "milplaintiff.cleaned.8.4.2020.csv")


plaintiff2
```

--**Context**: Berrada properties and its affiliates were responsible for at least 39% of all eviction filings since January 2019.


-- **Cleaned Milwaukee Plaintiffs since March 27 2020**
```{r}
top2 <- p2 %>% 
  filter(date > '2020-03-27') %>% 
  select(plaintiff_fixed, yearmon) %>% 
         group_by(plaintiff_fixed) %>% 
         count(plaintiff_fixed) %>% 
         filter(n > 5) %>% 
         ungroup() %>% 
         arrange(desc(n))
write.csv(top2, "cleaned_mil_evictors_mo_8_4_2020.csv")
top2
```

--**Context: Berrada was responsible for more than one out of every three evictions in Milwaukee since March 27, 2020.** Since the moratorium started in March 27, 2020, there have been 1,406 evictions. Berrada and its affiliated entities had 518 evictions, or 37% of the total.

```{r}
top2 <- p2 %>% 
  filter(date > '2020-03-27') %>% 
  select(plaintiff_fixed, yearmon) %>% 
         group_by(plaintiff_fixed, yearmon) %>% 
         count(plaintiff_fixed) %>% 
         filter(n > 5) %>% 
         ungroup() %>% 
         arrange(yearmon) %>% 
         arrange(desc(n))
write.csv(top, "cleaned_mil_evictors_mo_8_4_2020.csv")
```



--**Top with top 50 Results of Milwaukee Landlords**
```{r}
top2 %>% 
  top_n(50, n) %>% 
  kable() %>% 
  kable_styling("striped")

```



--**30**
