---
title: "Homeless Children in Arkansas"
author: "Rob Wells, Ph.D."
date: "3/28/2020"
output: 
  html_document:
      toc: true
      toc_float: true
      toc_depth: 4
---
 
## Reporting on Homelessness: Data Analysis for Journalists   

#### Jour 405v, Jour 5003, Spring 2020    

--------------------------------------------------------------------
# Part 1: Homeless Children Data
Load tidyverse, janitor, lubidate
```{r include=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```
Load Data
```{r include=FALSE}
ChildrenAR <- rio::import("https://github.com/profrobwells/HomelessSP2020/blob/master/Data/2018-2019-demographic-data-for-districts.xlsx?raw=true", which = "Data", skip=8)
```
Clean up column names to they are R friendly
```{r}
ChildrenAR <- janitor::clean_names(ChildrenAR)
```

- **Run a summary of the homeless data.** 

What is the median homeless percentage?   
What is the average (mean)?   
What is the maximum?   

```{r}
summary(ChildrenAR$district_percent_homeless)
```

- **Create a table with all school districts with more than 10% homeless**

HINT: SEE THE R_Intro_Jan15, the code for "Build a table with all counties above that median income"

```{r}
  ChildrenAR %>% 
  select(district_name, district_percent_homeless) %>% 
  filter(district_percent_homeless >= .10) %>% 
  arrange(desc(district_percent_homeless))

```
Create Children10 table, from code above, table with all school districts with more than 10% homeless
```{r}
Children10 <- ChildrenAR %>% 
  select(district_name, district_percent_homeless) %>% 
  filter(district_percent_homeless >= .10) %>% 
  arrange(desc(district_percent_homeless))
```

## Change Over Time

For your reference!
https://profrobwells.github.io/HomelessSP2020/SF_311_Calls_UofA.html

Import 2013-14 homeless data, call it ChildrenAR2013, clean the labels

```{r}
ChildrenAR2013 <- rio::import("https://github.com/profrobwells/HomelessSP2020/blob/master/Data/2013-2014-demographics-data-for-districts-2.xlsx?raw=true", which = "Data", skip=8)
# Clean up column names to they are R friendly
ChildrenAR2013 <- janitor::clean_names(ChildrenAR2013)
```

- **Create a table with just the district name and the homeless data for 2018-19**      

Call it table2018
```{r}
table2018 <- ChildrenAR %>% 
  select(district_name, district_percent_homeless, district_lea) %>% 
  arrange(desc(district_percent_homeless))
```

Rename a column name
```{r}
colnames(Children10)[2] <- "Pct_Homeless_2018"
#It takes the second column in Children10 and assigns the new name #"Pct_Homeless_2018"
names(Children10)
```
- **Rename the appropriate table2018 column to "Pct_Homeless_2018" ** 
```{r}
colnames(table2018)[2] <- "Pct_Homeless_2018"
```
- **Follow the same steps for the 2013 data:** 
Create table2013 
Rename the appropriate table2013 column to "Pct_Homeless_2013"
```{r}
table2013 <- ChildrenAR2013 %>% 
  select(district_name, district_percent_homeless,  district_lea) %>% 
  arrange(desc(district_percent_homeless))
#
colnames(table2013)[2] <- "Pct_Homeless_2013"
#
#convert all text to lowercase
#table2013$district_name <- tolower(table2013$district_name)
```
- **Joining Tables**   
```{r}
newtable <- table2018 %>% 
  inner_join(table2013, by=("district_lea"))
```
This matches the tables by name. They are joined by values in district_name   

- **How many records were left out?**    
8 Records.   
Use anti_join to determine the 2013 records not in 2018 
```{r}
table2013 %>% 
  anti_join(table2018, by=("district_lea"))
```

Making a table with the same thing above
```{r}
leftout <- table2013 %>% 
  anti_join(table2018, by=("district_lea"))
```

- **Math!**  

This creates a new column with the percentage point difference
```{r}
newtable <- newtable %>% 
  mutate(difference = Pct_Homeless_2018-Pct_Homeless_2013)
```

- **Produce a table with top 20 schools with the greatest increase in student homelessness since 2013, measured by percentage point difference**  

```{r}
top10worst <-  newtable %>% 
  top_n(20, difference) %>% 
  arrange(desc(difference))
```

# Basic Map Exercise
- **We will create a map of homeless children by schools in Arkansas**  

This lesson references the Machlis Ch 11 map exercise as a template. 
It also references Andrew Ba Tran's map tutorial
https://github.com/profrobwells/HomelessSP2020/blob/master/Readings/Tran%20-%20Geolocating.pdf

#### HERE WE GO!

Use ChildrenAR
Join ChildrenAR with addresses

Load Addresses
```{r}
School_Address <- rio::import("https://github.com/profrobwells/HomelessSP2020/blob/master/Data/SDI%20Districts%20Printed-02-26-2020-11_38_05.xlsx?raw=true", skip =1)
```

Convert School address LEA to numeric and match
```{r}
School_Address$LEA <- as.numeric(School_Address$LEA)
ChildrenAR$LEA <- ChildrenAR$district_lea
```
Now we join School_Address to ChildrenAR

```{r}
abc <- School_Address %>% 
  select(LEA, Name, Address) %>% 
  inner_join(ChildrenAR, by="LEA")
```

#QUESTION: How many records in new combined dataframe? What is the difference?

```{r}
abc_not <- School_Address %>% 
  select(LEA, Name, Address) %>% 
  anti_join(ChildrenAR, by="LEA")
```

Df with 261. which ones dropped? Lost two

NEXT, we geocode the addresses using mutate_geocode
This is a little involved... 

What I did:
?register_google
Google API Key https://www.youtube.com/watch?v=1JNwpp5L4vM
https://console.cloud.google.com/apis/credentials?folder=&organizationId=&project=phonic-botany-269820
register_google(key = "MY TOP SECRET KEY")
schools_geocoded <- mutate_geocode(abc, Address)
write.csv(schools_geocoded, "schools_geocoded.csv")

- **Shortcut!!!**

```{r include=FALSE}
schools_geocoded <- rio::import("https://raw.githubusercontent.com/profrobwells/HomelessSP2020/master/Data/schools_geocoded.csv")
```

Next: Simplify the school district names

```{r}
schools_geocoded$Name_Cleaned <- gsub("SCHOOL", "", schools_geocoded$Name)
```

```{r}
schools_geocoded$Name_Cleaned <- gsub("DISTRICT", "", schools_geocoded$Name)
```

```{r}
schools_geocoded$Name_Cleaned <- gsub("SCHOOL", "", schools_geocoded$Name_Cleaned) 
```
Load map town shapes for Arkansas with Tigris.

If you don't have tigris or ggplot2 or sf installed yet, uncomment and run the line below
```{r include=FALSE}
#install.packages("tigris", "sf", "ggplot2")
library(tigris) 
library(sf) 
library(ggplot2)
```
```{r}
# set sf option
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

And we finally load the school district boundary map
```{r include=FALSE}
# #Uncomment if you want to run this
# ar_schools <- school_districts("05")
```
