---
title: "Analysis of Washington County Jail Inmate Data"
author: "Michael Adkison and Rob Wells"
date: "3/31/2020"
output: 
  html_document:
      toc: true
      toc_float: true
      toc_depth: 4
---
 

#### Jour 405v, Jour 5003, Spring 2020    

#### Advanced Reporting and Data Analysis

#### Reporting on Homelessness in Northwest Arkansas

--------------------------------------------------------------------



```{r include=FALSE}
library(rio)
library(janitor)
library(tidyverse)
library(ggplot2)
theme_set(theme_classic())
```
### Importing Data
```{r}
Roster_March18 <- rio::import("https://docs.google.com/spreadsheets/d/1lsP37hPHXQ_KlBOsDJw5s7ehQlmeBqdX/edit#gid=814904708", which = "March18Roster")
Roster_March18 <- janitor::clean_names(Roster_March18)
```

Rename
```{r}
glimpse(Roster_March18)
```

```{r}
Homeless_March18 <- Roster_March18 %>%
  filter(homeless == "Yes")
```

```{r}
glimpse(Homeless_March18)
```

### Graphic for percentage of homelessness

```{r echo=TRUE}
df <- Roster_March18 %>%
  count(homeless)
df
```

```{r}
count.data <- df %>% 
  data.frame(prop = c(88.7, 11.3))
```

```{r}
count.data <- count.data %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)

theme_set(theme_classic(base_size = 18))

mycols <- c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF")

ggplot(count.data, aes(x = "", y = prop, fill = homeless)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
  geom_text(aes(y = lab.ypos, label = n), color = "white")+
  scale_fill_manual(values = mycols) +
  theme_void() +
  labs(title= "Homeless Detainees in Washington County Detention Center",
       subtitle="Data collected from detainee roster on March 28, 2020",
       caption= "Graphic by Michael Adkison",
       x= "Homeless Inmates",
       y= "Non-Homeless Inmates")
```


### Charges Facing Inmates

```{r}
ChargeStatistics <- rio::import("https://docs.google.com/spreadsheets/d/1lsP37hPHXQ_KlBOsDJw5s7ehQlmeBqdX/edit#gid=1846079051", which = "March18ChargeStats")
ChargeStatistics <- janitor::clean_names(ChargeStatistics)
```

Remove row 74 with totals
```{r}
ChargeStatistics <- ChargeStatistics[-c(74),] 
```
Summary of charges by category
```{r}
df2 <- ChargeStatistics %>% 
  summarize_if(is.numeric, sum, na.rm=TRUE)

df3 <- add_rownames(df2) %>% 
  gather(Charge, n, -rowname) %>% 
  spread(rowname, n) 

colnames(df3)[2] <- "Total"

df3$Charge <- gsub("felony_y_n", "Felony", df3$Charge)
df3$Charge <- gsub("drug_dwi", "Drug or DWI", df3$Charge)
df3$Charge <- gsub("administrative", "Administrative", df3$Charge)
df3$Charge <- gsub("assault_violence", "Assault of Violence", df3$Charge)
df3$Charge <- gsub("family_related", "Family Related", df3$Charge)
df3
```
Chart of charges

```{r}
ggplot(df3, aes(x = Charge, y = Total, fill= Total)) +
  geom_col(position = "dodge", show.legend=FALSE) +
  coord_flip() +   
  geom_text(aes(label= Total, x= Charge, y= Total), hjust=-1, vjust=1) +
    scale_y_continuous(limits=c(0, 75)) +
  labs(title= "Charges for Homeless Detainees",
       subtitle= "Washington Co. Sheriff Detainee Roster, 3/28/2020",
       caption= "Graphic by Michael Adkison",
       x= "",
       y= "Number of occurrences")
```

#### Wells Analysis
Compare Roster
https://drive.google.com/file/d/1lsP37hPHXQ_KlBOsDJw5s7ehQlmeBqdX/view?usp=sharing

    
    March 18 Roster = 637 detainees, 73 homeless or 11%
    March 27 Roster = 418 detainees, 42 homeless or 10%
    Overall Jail population reduced by 219 as March 27 from March 18. That accounts for 20 new detainees over the period.
    238 of the March 18 detainees were gone by March 27. Of this, 33 were homeless, or 14%.
    Records show 33 fewer homeless people in jail comparing the 399 matching records
    399 inmate names on March 27 remain, they match the March 18  roster
    20 new inmates since March 18
    3 of the new inmates are homeless
    
Clean names
```{r}
Roster_March18$name2 <- gsub(",", "_", Roster_March18$name_last_first)
Roster_March18$name2 <- gsub("\\s", "", Roster_March18$name2)
```
File exam of March 18

```{r}
#March 18 Roster: 637 detainees
nrow(Roster_March18)
```

```{r}
glimpse(Roster_March18)
```


```{r}
Roster_March27 <- rio::import("https://docs.google.com/spreadsheets/d/1lsP37hPHXQ_KlBOsDJw5s7ehQlmeBqdX/edit#gid=1284726909", which = "March27Roster")
Roster_March27 <- janitor::clean_names(Roster_March27)
#Clean Names
Roster_March27$name2 <- gsub(",", "_", Roster_March27$name_last_first)
Roster_March27$name2 <- gsub("\\s", "", Roster_March27$name2)
#March 27 Roster: 418 detainees
nrow(Roster_March27)
glimpse(Roster_March27)
```

Number homeless on March 18: 73
```{r}
bf <- Roster_March18 %>% 
  count(homeless) %>% 
  arrange(desc(n)) %>% 
  mutate(Pct = (100*(n/(sum(n)))),'%') 

bf$Pct <- round(bf$Pct, 2)
bf

```
Number homeless on March 27: 42
```{r}
wtf <- Roster_March27 %>% 
  count(homeless) %>% 
  arrange(desc(n)) %>% 
    mutate(Pct = (100*(n/(sum(n)))),'%') 

wtf$Pct <- round(wtf$Pct, 2)
nrow(wtf)
wtf
```
There are 31 fewer homeless people in jail since March 27. But this doesn't account for the number of homeless people released and the number of homeless added

Who remains - 399 match
```{r}
Remain <- Roster_March27 %>% 
  inner_join(Roster_March18, by="name2")
nrow(Remain)
```
```{r}
head(Remain)
```


Who was added - use antiJoin - shows 20 additions since March 18

```{r}
NewInmates <- Roster_March27 %>% 
  anti_join(Roster_March18, by="name2")
nrow(NewInmates)
```
```{r}
head(NewInmates)
```


Who departed - flip the antiJoin, start with March 18. 238 people

```{r}
Departed <- Roster_March18 %>% 
  anti_join(Roster_March27, by="name2")
head(Departed)
```

```{r}
glimpse(Departed)
```
### Chart who was released
33 Homeless people
```{r}
Departed %>% 
  count(homeless)
```



```{r}
Departed %>% 
  count(homeless) %>% 
ggplot() +
  geom_bar(mapping=aes(x= homeless, y= n, fill= n), stat = "identity", show.legend = FALSE) +
  coord_flip() +   
  geom_text(aes(label= n, x= homeless, y= n), hjust=-1, vjust=1) +
  scale_y_continuous(limits=c(0, 225)) +
  labs(title= "Housing Status for Released Detainees",
       subtitle= "Wash. Co. Sheriff Detainee Roster, 3/18 and 3/28/2020",
       caption= "Graphic by Michael Adkison",
       x= "Housing Status",
       y= "Number of released detainees")
```

